Compiler and Runtime Support for Functions-as-a-Service (FaaS)

### 2021-12-09 paper/paper.tex
[paper]: add rest of custom allocator subsection

\subsection{Profiler}
The profiler has the goal of understanding information about dynamic memory \ALLS\ and 
\DALLS\ during execution (for instance, the number of \ALLS\, \DWS\, common \ALL 
sizes, etc.) as well as typical information about critical paths or the "hotness" of functions and
loops at runtime that is often use to aid optimization [?]. Using this data, the profiler can inform the compiler during 
instrumentation in the second pass over the application (see \ref{ca}). It can also 
inform the user with a better estimate of memory usage for a future FaaS instance or 
function invocation.

@@

The allocator primarly takes the static analyses (see \ref{sa})
as input to replace as many \ALLS\ and \DALLS\ with invocations to the allocator. 
First, \ALLS\ with statically analyzable \SIZES\ are prime candidates to be 
\textit{pre-allocated} because all the information needed to allocate is known at
compile-time. These \ALLS\ are placed into a compiler-directed memory pool that is
allocated by the runtime in parallel to FaaS instance startup and before the entry-point function 
of the workload, $W$, begins. The size of the pool is the sum of all statically 
analyzable \SIZES\ of the selected \ALLS . As a result, the \ALL\ function calls are 
replaced by load of a compiler-determined offset of the base address of the pool.
An \ALL\ can only be selected for this transformation if it is not flagged by 
the escapes analysis (see the third analysis in \ref{sa}) \textit{and} the compiler 
can deduce the number of invocations of its parent function. Otherwise, the data
in memory or the pointer itself may be overwritten accidentally. The latter condition
can be alleviated with aggressive inlining and function versioning. An example of 
this transformation is shown in Figure~\ref{fig:pool}. There are a number of 
benefits to this approach: 1) alleviating function call overheads, 2) enabling 
prefetching of allocation pointer address, 3) hiding allocation overhead by performing 
it before $W$ even begins.

>>>>

\subsection{Profiler} \label{prof}
The profiler has the goal of understanding information about dynamic memory \ALLS\ and 
\DALLS\ during execution (for instance, the number of \ALLS\, \DWS\, common \ALL 
sizes, etc.) as well as typical information about critical paths or the "hotness" of functions and
loops at runtime that is often use to aid optimization [?]. Using this data, the profiler can inform the compiler during 
instrumentation in the second pass over the application (see \ref{ca}). It can also 
inform the user with a better estimate of memory usage for a future FaaS instance or 
function invocation.

@@

The allocator primarly takes the static analyses (see \ref{sa})
as input to replace as many \ALLS\ and \DALLS\ with invocations to the allocator. 
First, \ALLS\ with statically analyzable \SIZES\ are prime candidates to be 
\textit{pre-allocated} because all the information needed to allocate is known at
compile-time. These \ALLS\ are placed into a \textit{compiler-directed memory pool} that is
allocated by the runtime in parallel to FaaS instance startup and before the entry-point function 
of the workload, $W$, begins. The size of the pool is the sum of all statically 
analyzable \SIZES\ of the selected \ALLS . As a result, the \ALL\ function calls are 
replaced by load of a compiler-determined offset of the base address of the pool.
An \ALL\ can only be selected for this transformation if it is not flagged by 
the escapes analysis (see the third analysis in \ref{sa}) \textit{and} the compiler 
can deduce the number of invocations of its parent function or of the allocation
itself (if it belongs to a loop, for instance). Otherwise, the data
in memory or the pointer itself may be overwritten accidentally. The latter condition
can be alleviated with aggressive inlining and function versioning. An example of 
this transformation is shown in Figure~\ref{fig:pool}. There are a number of 
benefits to this approach: 1) alleviating function call overheads, 2) enabling 
prefetching of allocation pointer address, 3) hiding allocation overhead by performing 
it before $W$ even begins. This can boost performance and simplifies memory management
for FaaS instances.

Allocations that have statically analyzable size but could not be transformed to
use the compiler-directed memory pool, can still be transformed to invoke other
methods of the allocator. We motivate this by a simple example: a node 
to a linked list is allocated during an irregular while loop. These \ALLS\ most
likely belong to the same static IR instruction, but are invoked many times with
the same \SIZE\ during the loop. Consequently, we transform these \ALLS\ to 
invoke an allocation callback from a \textit{bump allocator} that has a block size equal to the 
\SIZE . The callback, \texttt{AllocateFromBump(blk\_size)}, returns a pointer from the
runtime's bump allocator dedicated to a block size of \texttt{blk\_size}. The bump allocator
for \texttt{blk\_size} is initialized in parallel to FaaS instance startup via 
a compiler-injected callback, since \texttt{blk\_size} is already known at compile-time.
All bump allocators are stored by their block sizes in the runtime.

Finally, \ALLS\ that may be invoked many times (as in the example), but have 
variable-dependent \SIZES\ (i.e. the size is not a constant integer) are transformed to 
invoke a callback, \texttt{AllocateFromBumpWithRuntimeInit(blk\_size)}, that performs 
bump allocator initialization during runtime if the allocator for \texttt{blk\_size} is not
yet initialized. \ALLS\ that fit the criteria for this transformation include ones
in loops with loop-invariant and variable-dependent \SIZES .

It is important to note that the profiler (see \ref{prof}) can aid these transformations.
For instance identifying hot functions can expand the number of \ALLS\ that could
be transformed to invoke \texttt{AllocateFromBumpWithRuntimeInit}, since those static
\ALLS\ would be executed many times. Moreover, each bump allocator starts with a pool of 
a default number of of entries to allocate from before expanding to more pools. Using
profile information, the number of entries could be tuned for each specific bump allocator.
